---
title: "Statistics Pipeline"
output:
  html_notebook:
    toc: yes
    toc_float:
      smooth_scroll: yes
      collapsed: no
  html_document:
    toc: yes
    df_print: paged
date: "Last edited: `r format(Sys.time(), '%Y-%m-%d')`"
---

```{r dataframe, include=F}
Target <- rep("hIFIT1", times = 12)
Condition <- factor(rep(c("Mock", "cbRSV dNS1 0.001-24","cbRSV dNS2 0.01-24", "cbRSV dNS1/2 0.001-24"), each = 3), levels = c("Mock", "cbRSV dNS1 0.001-24","cbRSV dNS2 0.01-24", "cbRSV dNS1/2 0.001-24"))
Value_norm <- c(
  0.470688086, 1.584484165, 0.944827749, 81.20096936, 24.32041674, 75.01759177, 24.90206996, 9.923147271,19.3690698, 278.2460496, 292.1552481, 24.35566374
)
a549_multiple <- data.frame(Target, Condition, Value_norm)
a549_single <- a549_multiple %>% filter(Condition %in% c("Mock", "cbRSV dNS1 0.001-24"))
```

```{r tidyverse_load, include=F}
library(tidyverse)
```

# Testing Statistical Assumptions

```{r Dataset}
# Replace 'a549_multiple' and 'a549_single' in the workbook
```

## Visual Tests

#### Normal Distribution by Boxplot

```{r boxplot}
boxplot(Value_norm~Condition, a549_multiple)
```

#### Testing equality of variance assumptions

```{r testing_assumptions}
plot(lm(Value_norm~Condition, a549_multiple))
```

**1st and the last plots:** we want symetrical data about the 0 horizontal line

**2nd plot:** we want residual points to be as close to the predicted line as possible

**3rd plot:** we want for red line to be approx. horizontal

## Mathematical tests

### Testing Normality

#### Test Normality (of distribution) per treatment group

**p value \> 0.05 means normal distribution**

```{r shapiro.test_single}
shapiro.test(a549_multiple$Value_norm[1:3])
shapiro.test(a549_multiple$Value_norm[4:6])
shapiro.test(a549_multiple$Value_norm[7:9])
shapiro.test(a549_multiple$Value_norm[10:12])
```

#### Test Normality (of distribution) for the whole dataset

Visual assessment where you want the residual points to be as close line as possible.

```{r shapiro.test_total}
plot(residuals(lm(Value_norm~Condition, a549_multiple)))
shapiro.test(residuals(lm(Value_norm~Condition, a549_multiple)))
```

If all groups and combined analysis evaluate to normal distribution then its normal. If one group is non normal by itself but combined dataset evaluates to normal distribution usually you can assume its normal distribution.

### Testing the Equality of Variance

#### For **NORMAL** Distribution

**p value \> 0.05 means equal variance**

```{r bartlett.test}
bartlett.test(Value_norm~Condition, a549_multiple)
```

#### For **NON NORMAL** Distribution

**p value \> 0.05 means equal variance**

```{r leveneTest}
library(car, quietly = T)
leveneTest(Value_norm~Condition, a549_multiple)
```

# Obtaining Statistical Parameters

## For **Normal Distribution** and **Equal Variance**

### Multiple Comparisons

```{r anova}
library(DTK, quietly = T)
anova(lm(Value_norm~Condition, a549_multiple))
```

Anova finds there is any significant difference across the whole dataset. If the p-value (Pr(>F) is **ABOVE** 0.05 the analysis should be stopped here!

Important parameters: F value and Pr(>F) (include in reports)

```{r tukey}
TukeyHSD(aov(Value_norm~Condition, a549_multiple))
```

P adj are the individual p values between group combinations.

### Single Comparison

```{r paired_t_test}
t.test(Value_norm~Condition, 
       data=a549_single, 
       alternative='two.sided',
       var.equal=T)
```

Important parameters: t and p-value (include in reports)

## For **Non Normal Distribution** but **Equal Variance**

### Multiple Comparisons

```{r kruskal.test}
library(dunn.test, quietly = T)
dunn.test(a549_multiple$Value_norm, 
          a549_multiple$Condition, 
          altp=T,
          list=T)
```

Kruskal test finds there is any significant difference across the whole dataset. If the p-value is **ABOVE** 0.05 the analysis should be stopped here without comparing groups!

Important parameters: chi-squared and p-value (include in reports)

The dunn test displays both comparison matrix and comparison list of tested groups. P values are the individual p values between group combinations.

### Simple Comparison

```{r paired_t_test_2}
t.test(Value_norm~Condition, 
       data= a549_single, 
       alternative='two.sided',
       var.equal=T)
```

Important parameters: t and p-value (include in reports)

## For **Normal Distribution** but **Non Equal Variance**

### Multiple Comparisons

```{r one_way_test}
oneway.test(Value_norm~Condition, 
            a549_multiple, 
            var.equal = F)
```

One-way analysis of means finds there is any significant difference across the whole dataset. If the p-value is **ABOVE** 0.05 the analysis should be stopped here!

Important parameters: F value and Pr(>F) (include in reports)

```{r games_howell}
library(rstatix, quietly = T)
library(dplyr, quietly = T)
a549_multiple %>% games_howell_test(Value_norm~Condition)
```

P adj are the individual p values between group combinations.

### Single Comparison

```{r Welch_Two_Sample_t-test}
t.test(Value_norm~Condition, 
       data=a549_single, 
       alternative='two.sided',
       var.equal=F)
```

Important parameters: t and p-value (include in reports)