---
title: "Example of qPCR Pipeline Usage"
subtitle: "Copy number GAPDH normalised dataset"
author: "Michal Varga"
date: "Last edited: `r format(Sys.time(), '%Y-%m-%d')`"
output: rmarkdown::github_document
params:
    target_gene: bIFIT1
    primer_set: PS11
    cell_line: MDBK
    housekeeping_gene: bGAPDH
    conditions: !r c('Mock', 'hRSV 1-24', 'bRSV dSH 1-24')
    colors: !r c('#999999', '#ffc034', '#9a6a00')
    statistics: "Normal distribution and equal variance"
    pvals: !r c(0.9935156, 0.1817141)
    range: 14
    breaks: 2
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r reading_ext_script, include=FALSE, cache=FALSE, eval=FALSE}
knitr::read_chunk("../Report/Templates/external_chunk.R")
```

```{r knitr_libraries, include=F}
library(knitr)
library(kableExtra)
library(rmarkdown)
```

# Introduction

Purpose: 

based on scripts:

-   standard curve
-   control gene Ct based
-   Copy number GAPDH normalised dataset

parameters incuded in script:

-   `target_gene:` bIFIT1
-   `housekeeping_gene:` bGAPDH
-   `conditions:` !r c('Mock', 'phRSV 1-24','bIFNa 5-6')
-   `colors:` !r c('#999999', '#ffc034', '#9a6a00')
-   `statistics:` "Normal distribution and equal variance"
-   `pvals:` !r c(0.9935156, 0.1817141)
-   `range:` 14
-   `breaks:` 2

Link to stats pipeline?

## Libraries (maybe put them when they are used?)

```{r required_libraries, warning=FALSE, message=FALSE}
library(tidyverse)
library(data.table)
library(ggsignif)
library(scales)
```

# 1. Standard Curve

## Introduction

Some text about what the fuck

## Data Loading

Load `standard_curves_data.csv` from `/Data` folder, and filter the target gene and primer set used based on the declared parameters in YAML.

```{r data_loading_standard_curve}
standard_curve_data <- fread('../Data/standard_curves_data.csv') %>%
  filter(Target == params$target_gene,
         Primer_set == params$primer_set)
```

```{r data_loading_standard_curve_display, echo=FALSE, warning=FALSE}
standard_curve_data |> kable()
```

## Linear Model Production

### Construct Linear Model
```{r standard_curve_model}
model_standard_curve <- lm(log10(Copy_number)~Ct, data = standard_curve_data)

```

```{r standard_curve_model_display, echo=FALSE, warning=FALSE}
lm(log10(Copy_number)~Ct, data = standard_curve_data)
```

### Construct a Prediction Based on the Linear Model

```{r standard_curve_model_prediction}
prediction_model_standard_curve <- 10^predict(
  model_standard_curve, 
  interval = 'prediction'
  )

```
```{r standard_curve_model_prediction_display, echo=FALSE, warning=FALSE}
10^predict(
  model_standard_curve, 
  interval = 'prediction'
  ) |> kable()
```


```{r standard_curve_dataset_bind}
data_combined_standard_curve <- cbind(
  standard_curve_data,
  prediction_model_standard_curve
  ) # save it to a variabel

```
```{r standard_curve_dataset_bind_display, echo=FALSE, warning=FALSE}
cbind(
  standard_curve_data,
  prediction_model_standard_curve
  ) |> kable()
```


### Calculate Amplification Efficiency

Some text about what the fuck are we doing

Put amplification efficiency equation

```{math}
\text{Amplification Efficiency} = 10^{-1/\text{slope}}-1
```

```{r standard_curve_amplification_efficiency}
efficiency_standard_curve <- paste(
  round(
    (10^(-1/ lm(Ct~log10(Copy_number), 
                data = standard_curve_data)[[1]][2]) -1)*100, 
    digits = 2), 
  '% Amplification Efficiency', 
  sep = ''
  ) # save it as a variable

```

```{r standard_curve_amplification_efficiency_display, echo=FALSE, warning=FALSE}
paste(
  round(
    (10^(-1/ lm(Ct~log10(Copy_number), 
                data = standard_curve_data)[[1]][2]) -1)*100, 
    digits = 2), 
  '% Amplification Efficiency', 
  sep = ''
  )

```

## Figure

### Constructing Figure

```{r standard_curve_figure_parameters}
plot_title_standardCurve <- paste(params$target_gene, 
                    '-', 
                    params$primer_set,
                    'Standard Curve', 
                    sep = " ")
x_annotation_position_standardCurve <- 1000000
y_annotation_position_standardCurve <- 37.5
top_range_standardCurve <- 40
y_axis_title_standardCurve <- 'Cycle Threshold'
x_axis_title_standardCurve <- 'Copy Number'
```

```{r standard_curve_figure_initial_ggplot}
plot_standard_curve <- ggplot(
  data = data_combined_standard_curve,
  mapping = aes(x = Copy_number,
                y = Ct)) +
  geom_point() +
  stat_smooth(method = lm) +
  labs(
    title = plot_title_standardCurve,
    y = x_axis_title_standardCurve,
    x = y_axis_title_standardCurve
  )
```

```{r standard_curve_figure_y_axis}
plot_standard_curve <- plot_standard_curve +
  scale_y_continuous(
    breaks = seq(from = 0, 
                 to = top_range_standardCurve, 
                 by = 10),
    limits = c(0,top_range_standardCurve)
  )
```

```{r standard_curve_figure_log_scale}
plot_standard_curve <- plot_standard_curve +
  scale_x_log10(
    labels = trans_format("log10", 
                          math_format(10^.x)),
    breaks = trans_breaks("log10", 
                          function(x) 10^x)
  ) +
  annotation_logticks(sides='b')
```

```{r standard_curve_figure_theme}
plot_standard_curve <- plot_standard_curve +
  theme_bw() +
  theme(
  plot.title = element_text(
    size = 15, 
    face = 'bold', 
    margin = margin(8, 0, 8, 0), 
    hjust = 0.5
  ),
  axis.text.y = element_text(
    angle = 0, 
    size = 9, 
    vjust = 0.5),
  axis.text.x.bottom = element_text(
    angle = 0, 
    size = 9, 
    vjust = 0.5),
  axis.title.x = element_text(
    size = 12, 
    face = 'bold', 
    vjust = -0.5, 
    margin = margin(0, 8, 0, 0)),
  axis.title.y = element_text(
    size = 12, 
    face='bold', 
    vjust=-0.5, 
    margin = margin(0, 8, 0, 0)),
  aspect.ratio = 1/2
)
```

```{r standard_curve_figure_annotation}
plot_standard_curve <- plot_standard_curve +
  annotate('text',
           y = y_annotation_position_standardCurve, 
           x = x_annotation_position_standardCurve, 
           label = efficiency_standard_curve, 
           size = 5)
```


```{r standard_curve_plot_display, , echo=FALSE, warning=FALSE}
plot_standard_curve
```

### Saving Figure

```{r standard_curve_figure_save}
ggsave(filename = paste(
  paste("plot_standard_curve", 
                    params$target_gene, 
                    params$primer_set,
                    sep = "_"), '.png', sep = ''), 
       plot = plot_standard_curve, 
       device = 'png', 
       path = '../Figures', 
       dpi = 600, 
       height = 16, 
       width = 20, 
       units = 'cm')
```


# 2. Housekeeping Gene Control

## Introduction

Some text about what the fuck

## Data Loading

Load `ct_data.csv` from `/Data` folder, and filter the target gene and primer set used based on the declared parameters in YAML.


```{r data_loading_housekeeping}
housekeeping_gene_data <- fread('../Data/ct_data.csv') %>% 
  filter(Target == params$housekeeping_gene, 
         Cell_line == params$cell_line,
         Condition %in% params$conditions)
```

```{r data_loading_housekeeping_display, echo=FALSE, warning=FALSE}
housekeeping_gene_data |> kable()
```

## Processing Data

wriet about what the fuck is happening

equation for ddCt 

```{math}
\text{Relative Quantification} = 2^{\Updelta\Updelta \text{Ct}}
```
```{math}
\Updelta\Updelta \text{Ct} = \Updelta \text{Ct}_{\text{Test Samples}}-\Updelta \text{Ct}_{\text{Calibrator Samples}}
```
```{math}
\Updelta \text{Ct}_{\text{Test Samples}} = \text{Ct}_{\text{Target Gene in Tests}}-\text{Ct}_{\text{Reference Gene in Tests}}
```
```{math}
\Updelta \text{Ct}_{\text{Calibrator Samples}} = \text{Ct}_{\text{Target Gene in Calibrator}}-\text{Ct}_{\text{Reference Gene in Calibrator}}
```


```{r housekeeping_gene_factors}
housekeeping_gene_data <- housekeeping_gene_data |> mutate(
  control_mean = mean(
    Ct[Condition == params$conditions[1]],
    na.rm = T),
  log2_dCt = 2^ (- (Ct - control_mean)),
  control_mean_log = mean(
    log2_dCt[Condition == params$conditions[1]],
    na.rm = T),
  Value_norm = log2_dCt / control_mean_log
)
```

```{r housekeeping_gene_factors_display, echo=FALSE, warning=FALSE}
housekeeping_gene_data |> kable()
```

## Factorisation

REMANE ME!!!

```{r factroisation}
housekeeping_gene_data <- aggregate(x=housekeeping_gene_data, 
                                    by=list(housekeeping_gene_data$Condition),
                                    FUN = mean)
housekeeping_gene_data <- housekeeping_gene_data %>%
  arrange(match(Group.1, params$conditions))

housekeeping_factor_vector <- rep(housekeeping_gene_data$Value_norm, each=3)
```


# 3. Factorised Copy Number Extrapolation

## Introduction

Some text about what the fuck

## Data loading and filtering

Load `ct_data.csv` from `/Data` folder, and filter the target gene and primer set used based on the declared parameters in YAML.

```{r data_loading}
target_data <- fread('../Data/copy_number_extrapolation_data.csv') %>%
  filter(Target == params$target_gene,
         Cell_line == params$cell_line,
         Condition %in% params$conditions)
```

```{r data_loading_display, echo=FALSE, warning=FALSE}
target_data |> kable()
```

## Copy number extrapolation

```{r copy_number_extrapolation}
target_data_modelled <- target_data |> 
  mutate(
    Copy_number = 10^predict(model_standard_curve, 
                             newdata = target_data))
```

```{r copy_number_extrapolation_display, echo=FALSE, warning=FALSE}
target_data_modelled |> kable()
```


## Factorisation based on housekeeping gene levels

wriet about what the fuck is happening

equation for ddCt 

```{math}
\text{Relative Quantification} = 2^{\Updelta\Updelta \text{Ct}}
```
```{math}
\Updelta\Updelta \text{Ct} = \Updelta \text{Ct}_{\text{Test Samples}}-\Updelta \text{Ct}_{\text{Calibrator Samples}}
```
```{math}
\Updelta \text{Ct}_{\text{Test Samples}} = \text{Ct}_{\text{Target Gene in Tests}}-\text{Ct}_{\text{Reference Gene in Tests}}
```
```{math}
\Updelta \text{Ct}_{\text{Calibrator Samples}} = \text{Ct}_{\text{Target Gene in Calibrator}}-\text{Ct}_{\text{Reference Gene in Calibrator}}
```


## Statistics


## Plotting


# Libraries and Data Loading

the libraries needed and why

```{r, plot_theme, eval=FALSE}
```

# Data Manipulation

## Linear Model Creation

```{r linear_model_creation, eval=FALSE}
model_lmscdata <- lm(log10(Copy_number)~Ct,  data = scdata)
```

## Linear Model Figure

```{r, lm_figure_loading, echo=F, eval=FALSE}
```

```{r lm_figure, eval=FALSE}
plot_scdata
```

## Housekeeping Log2 Data Generation

```{r housekeeping_data, eval=FALSE}
housekeeping_gene_data <-  housekeeping_gene_data %>% 
  mutate(control_mean = mean(Ct[Condition == params$conditions[1]], na.rm = T),
        log2_dCt = 2^ (- (Ct - control_mean)),
        control_mean_log = mean(log2_dCt[Condition == params$conditions[1]], na.rm = T),
        Value_norm = log2_dCt / control_mean_log) %>%
  group_by(Condition) %>%
  mutate(mean_Ct = mean(Ct)) %>%
  ungroup() %>%
  arrange(match(Condition, params$conditions))
```

## Extrapolation From Standard Curve and Factorisation by Housekeeping Gene Data

```{r data_manipulation, eval=FALSE}
target_data <- target_data %>% 
  arrange(match(Condition, params$conditions)) %>%
  mutate(Copy_number = 10^predict(model_lmscdata, newdata = target_data),
         Factor = housekeeping_gene_data$mean_Ct,
         Copy_number_mod = Copy_number / Factor,
         Control_mean = mean(Copy_number_mod[Condition == params$conditions[1]], na.rm = T),
         Value_norm = Copy_number_mod / Control_mean)
```

```{r kable_final_table, eval=FALSE}
kable(target_data, caption = "Final Table")
```

## Dataset Save

```{r dataset_save, eval=FALSE}
print("SAVE THE DATAAA")
```

Data was saved to **HELL**

# Statistics

## `r params$statistic`

```{r statistics, eval=FALSE}
library(DTK, quietly = T)
anova(lm(Value_norm~Condition, target_data))
TukeyHSD(aov(Value_norm~Condition, target_data))
```

# Data Visualisation

```{r, final_plotting, eval=FALSE}
```

```{r significance_bars, eval=FALSE}
plot_target_data +
  geom_signif(
    comparisons = list(c(
      params$conditions[1], 
      params$conditions[2])), 
    annotation = params$pvals[1], 
    y_position = 0.93*params$range - 1*(params$range*0.075), 
    tip_length = 0, vjust= -0.2, size = 0.7,
    textsize = textsize_values[1]) +
              
  geom_signif(
    comparisons = list(c(
      params$conditions[1], 
      params$conditions[3])), 
    annotation = params$pvals[2], 
    y_position = 0.93*params$range - 0*(params$range*0.075), 
    tip_length = 0, vjust= -0.2, size = 0.7, 
    textsize = textsize_values[2])
```

## Figure Save

```{r figure_save, eval=FALSE}
print("SAVE THE FIGUREEEEE")
```

Figure was saved to **HELL**
